%LaTeX example file 2

%basic text editing
%this is also a skeleton for a two-up handout

\documentclass[10pt, twoside]{article}

\usepackage{tipa}
\usepackage{geometry}
\usepackage[normalem]{ulem}
%usepackage{doublespace}
%\usepackage{times}
%usepackage{palatino}

\geometry{margin=1in}

\title{Basic Font Commands}
\author{Charlie Kelly\\UC-Santa Cruz}


\begin{document}
	
	\maketitle
	
		\section{Basic Text Formatting}
	
	Why don't I strap on my \textsc{job helmet} and squeeze down into a \textbf{job cannon} and fire off into \emph{Jobland} where \textit{jobs} grow on \uline{little jobbies.}
	
	The {\bf tipa} package can /\textipa{hElp wiT fonEtIks}/.
	
	\textsl{I'm cracking eggs of wisdom!}
	
	\uuline{Look at the door, dude.} \uwave{You see that door right there?} \sout{The one marked pirate?}\footnote{You think a pirate lives in there?} (This requires the package \texttt{ulem}).
	
	There are a couple of useful text diacritics: \'{e} \`{a} \"{o} \~{n} \^{i} \textsubdot{t} \={o}.
	
	Superscript and subscript require math mode (to be discussed later)
	
	\section{Sizes}
	
	Font sizes are changed by setting the font size with a command; you must then set it back with another text font command.
	
	\tiny
	
	They took you nightman and you don't belong to them.
	
	\scriptsize
	
	Wait wait wait! What are you gonna do? Punch him in the face? Throw him? Maybe work the body a little?
	
	\small
	
	I'm in love with a man\ldots a man named God.
	
	\normalsize
	
	This is the normal size for the document (10-12pt, as defined in the call to \texttt{documentclass}).
	
	\large
	
	Also, I am familiar with carpentry and I don't know who my father is. So, am I the messiah?
	
	\Large
	
	That's called flip-flopping, Mac! That's what Democrats do.
	
	\LARGE
	
	I become the spirit of the Nightman.
	
	\huge
	
	But I'm really Dayman.
	
	\Huge
	
	This is as big as it gets.
	
\end{document}