%LaTeX Example file 3

%This document contains some basic document structure examples
%the itemize and enumerate environment
%basic refernce counting

\documentclass[12pt, letterpaper]{article}

\usepackage{tipa}

\title{Structuring Your Document}
\author{Marshall Eriksen\\Columbia University}
\date{}

\begin{document}
	
	\maketitle
	
	\section{Sections}
	\label{sec}
	
	Look at us, riding around in a limo, eating hot dogs\ldots it's like we're the president.
	
	\subsection{Subsections}
	\label{subsec}

	\subsubsection{Subsubsections}
	\label{subsubsec}
	
	My stomach was like, ``Hey bro, I don't know what you're eating 'cause I don't have any eyes but it's basically awesome so keep sending it down Gullet Alley." "But don't do quotes like this.""
	
	\paragraph{Paragraph} ``She's Like the Wind'' has been stuck in my brain for about 40 hours. I just got it out; now it's back in.
	
	\subparagraph{Subparagraph} This is as deeply-embedded as it gets.
	
	\clearpage
	
	\section{Itemize and Enumerate}
	
	\begin{itemize}
		\item Why are Ewoks so awesome?
		\begin{itemize}
			\item The Rebellion would have totally failed without Ewoks.
			\item[\textbf{Penguins}:] are like Ewoks but without the fur.
		\end{itemize}
		\item The Loch Ness monster:
		\begin{itemize}
			\item Is totally real
			\begin{itemize}
				\item exists
				\begin{itemize}
					\item is awesome.
				\end{itemize}
			\end{itemize}
		\end{itemize}
	\end{itemize}
	
	\begin{enumerate}
		\item Enumerate
		\begin{enumerate}
			\item Is useful for making outlines
			\item and other things
		\end{enumerate}
	\end{enumerate}
	
	\section*{References}
	
		You can refer to a reference with \texttt{ref\{\}}, but then you must run \LaTeX twice to have it track the references for you. \S{\ref{sec}} was labeled, as was its daughter \S\S\ref{subsec}.

\end{document}