\documentclass[12pt, letterpaper]{article}

\usepackage{graphicx}
\usepackage[letterpaper, margin=1in]{geometry}
\usepackage{jtree}
\usepackage{qtree}
\usepackage{gb4e}
\usepackage{tipa}
\usepackage{pstricks}
\usepackage{link}
\usepackage{linguex}


\begin{document}

% linguex examples:

\ex. Bill thinks that they like each other.

\ex. \a. *They think that Bill likes each other.
\b. *They think that each other please Bill.

\ex. \label{gen:aae-rescue}AAE Rescue Strategies:
\a. No agreement (Inuit)
\b. Default agreement (Italian)\footnote{I have omitted for space reasons the cases of Georgian and Albanian, which \cite{woolford99} shows also employ this strategy with much more grammatical results than in Italian.}
\c. Anaphoric agreement (Swahili)
\d. Poessessive reflexives (Selayarese)
\e. Detransitivizing Morphology (French)


\ex. Postverbal Theme Complementary Distribution:
\ag. \label{ex:aae-italian2}*{\textipa{a loro}} \textipa{interessano} \textipa{solo} {\textipa{se stessi}}\\
{to them(.\textsc{dat})} interest.\textsc{3.pl} only {themselves(.\textsc{nom})}\\
``They are interested only in themselves.'' \hfill \citep[p.33]{rizzi90a}
\bg. {\textipa{a me}} \textipa{interessano} \textipa{solo} \textipa{loro}\\
{to me(.\textsc{dat})} interest.\textsc{3.pl} only they(.\textsc{nom})\\
``I am interested only in them.'' \hfill \citep[p.32]{rizzi90a}

\ex. Nez Perce Reflexive Prefix Series \citep[p.40]{rude85}:\\
\begin{tabular}[h!]{lll}
  &\textsc{sg}&\textsc{pl}\\
\textsc{1}&\textipa{'in\'{e}e-}&\textipa{nem\'{e}e-}\\
\textsc{2}&\textipa{'im\'{e}e-}&\textipa{'imem\'{e}e-}\\
\textsc{3}&\textipa{'ipn\'{e}e-}&\textipa{'imem\'{e}e-}
\end{tabular}

%% Linuex NB: \Next is the next example, \NNext the example after that. Similarly with \Last and \LLast


% gb4e examples:


\begin{exe}
  \ex Regular \emph{+ee} Affixation:
  \begin{xlist}
    \ex present $\rightarrow$ present\emph{ee}
    \ex employ $\rightarrow$ employ\emph{ee}
    \ex pay $\rightarrow$ pay\emph{ee}
  \end{xlist}
\end{exe}

\begin{exe}
  \ex \begin{xlist}
   \ex I am strongly inclined to believe it.
   \ex I am more strongly inclined to believe it.
   \ex I am stronger inclined to believe it.
  \end{xlist}
  \ex \begin{xlist}
    \ex He ran quickly (*quick).
    \ex He ran more quickly.
    \ex He ran quicker.
  \end{xlist}
  \ex \begin{xlist}
    \ex He spoke softly (*soft).
    \ex He spoke more softly.
    \ex He spoke softer.
  \end{xlist}
\end{exe}

\begin{exe}
  \ex
  \gll [[root + A]_X + B]_Y\\
  1 { } 2 { } 3 $\rightarrow$ 1 $\phi$ 3\\
  where X an Y are major lexical categories
\end{exe}

\begin{exe}
  \ex Adverb \emph{+er} Truncation:
  \gll {C_0VC_0 +} {ly +} {er_{Adv}}\\
  1 2 3 $\rightarrow$ 1 $\phi$ 3\\
\end{exe}

\end{document}